\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2012-10-27 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
renamed decompose_mesh_SCOTCH to decompose_mesh because METIS is now supported as well, in addition to SCOTCH (and ZOLTAN may also be added in the future).

The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
It will also use the PT-SCOTCH parallel library for mesh partitioning.\\

@@

The SPECFEM3D Cartesian software package relies on the SCOTCH library to partition meshes created with CUBIT.
Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.

@@

These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh\_SCOTCH}
to load-balance the mesh. Please see the next section for further details.

@@

\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh\_SCOTCH}}

@@

An efficient partitioning is important because it leverages the overall running time of the application.
It amounts to balance the number of elements in each slice while minimizing the communication costs
resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh\_SCOTCH} depends on the SCOTCH library \citep{PeRo96},
which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
\\

In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
During the configuration process, the script tries to find existing SCOTCH installations.
In case your system has no pre-existing SCOTCH installation,
we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
in directory
\texttt{src/decompose\_mesh\_SCOTCH/scotch\_5.1.10b}.
This version gets bundled with the compilation of the SPECFEM3D Cartesian package if no libraries could have been found.
If this automatic compilation of the SCOTCH libraries fails,
please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
In case you want to use a pre-existing installation, make sure you have correctly specified the path
of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
\\

@@

\caption{Example of a mesh partitioning onto four cores. Each single core partition is colored differently. The executable \texttt{xdecompose\_mesh\_SCOTCH} can equally distribute the mesh on any arbitrary number of cores.
Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}

@@

When you are ready to compile,
in the main directory type
\begin{lyxcode}
{\small make xdecompose\_mesh\_SCOTCH}{\small \par}
\end{lyxcode}
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh\_SCOTCH} should be produced.

The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
the synopsis is:
\begin{lyxcode}
./bin/xdecompose\_mesh\_SCOTCH  nparts  input\_directory output\_directory
\end{lyxcode}
where
\begin{itemize}
     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{MESH/}, and
     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition. These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}. Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
\end{itemize}
Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.

@@

After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh\_SCOTCH}, the next step in the workflow is to compile
\texttt{xgenerate\_}~\\
\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.

@@

the partitioner, i.e. from \texttt{xdecompose\_mesh\_SCOTCH} or \texttt{xmeshfem3D}.

>>>>

The next release of the code will include support Convolution or Auxiliary Differential Equation Perfectly Matched absorbing Layers
(C-PML or ADE-PML) \citep{KoMa07,MaKoEz08,MaKoGe08,MaKo09,MaKoGeBr10}.
It will also use the PT-SCOTCH parallel and threaded version of SCOTCH for mesh partitioning instead of the serial version.\\

@@

Note that we use CUBIT to create meshes of hexahedra, but other packages can be used as well, for instance GiD from \url{http://gid.cimne.upc.es} or Gmsh from \url{http://geuz.org/gmsh} \citep{GeRe09}. Even mesh creation packages that generate tetrahedra, for instance TetGen from \url{http://tetgen.berlios.de}, can be used because each tetrahedron can then easily be decomposed into four hexahedra as shown in the picture of the TetGen logo at \url{http://tetgen.berlios.de/figs/Delaunay-Voronoi-3D.gif}; while this approach does not generate hexahedra of optimal quality, it can ease mesh creation in some situations and it has been shown that the spectral-element method can very accurately handle distorted mesh elements \citep{OlSe11}.

The SPECFEM3D Cartesian software package relies on the SCOTCH library to partition meshes created with CUBIT.
METIS \citep{KaKu98a,KaKu98c,KaKu98b} can also be used instead of SCOTCH if you prefer, by editing file src/decompose\_mesh/decompose\_mesh.F90 and uncommenting the flag USE\_METIS\_INSTEAD\_OF\_SCOTCH. You will also then need to install and compile Metis version 4.0 (do *NOT* install Metis version 5.0, which has incompatible function calls) and edit \texttt{src/decompose\_mesh/Makefile.in} and uncomment the METIS link flag in that file before running \texttt{configure}.\\

@@

These mesh files are needed as input files for the partitioner  \texttt{xdecompose\_mesh}
to load-balance the mesh. Please see the next section for further details.

@@

\subsection{Partitioning the Mesh with \texttt{xdecompose\_mesh}}

@@

An efficient partitioning is important because it leverages the overall running time of the application.
It amounts to balance the number of elements in each slice while minimizing the communication costs
resulting from the placement of adjacent elements on different processors. \texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
which provides efficent static mapping, graph and mesh partitioning routines. SCOTCH is a free software package developed by
Fran\c {c}ois Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable from the web page \url{https://gforge.inria.fr/projects/scotch/}.
\\

In most cases, the configuration with \texttt{./configure~FC=ifort} should be sufficient.
During the configuration process, the script tries to find existing SCOTCH installations.
In case your system has no pre-existing SCOTCH installation,
we provide the source code of SCOTCH, which is released open source under the French CeCILL-C version 1 license,
in directory
\texttt{src/decompose\_mesh/scotch\_5.1.12b}.
This version gets bundled with the compilation of the SPECFEM3D Cartesian package if no libraries could have been found.
If this automatic compilation of the SCOTCH libraries fails,
please refer to file INSTALL.txt in that directory to see further details how to compile it on your system.
In case you want to use a pre-existing installation, make sure you have correctly specified the path
of the SCOTCH library when using the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure} script.
In the future you should be able to find more recent versions at  \url{http://www.labri.fr/perso/pelegrin/scotch/scotch\_en.html} .
\\

@@

\caption{Example of a mesh partitioning onto four cores. Each single core partition is colored differently. The executable \texttt{xdecompose\_mesh} can equally distribute the mesh on any arbitrary number of cores.
Domain decomposition is explained in detail in \cite{MaKoBlLe08}, and excellent scaling up to 150,000 processor cores in shown for instance in
\cite{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,KoGoErMi10,Kom11}.}

@@

When you are ready to compile,
in the main directory type
\begin{lyxcode}
{\small make xdecompose\_mesh}{\small \par}
\end{lyxcode}
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh} should be produced.

The partitioning is done in serial for now (in the next release we will provide a parallel version of that code),
the synopsis is:
\begin{lyxcode}
./bin/xdecompose\_mesh  nparts  input\_directory output\_directory
\end{lyxcode}
where
\begin{itemize}
     \item \texttt{nparts} is the number of partitions, i.e., the number of cores for the parallel simulations,
     \item \texttt{input\_directory} is the directory which holds all the files generated by the Python script \texttt{cubit2specfem3d.py} explained in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{MESH/}, and
     \item \texttt{output\_directory} is the directory for the output of this partitioner which stores ACII-format files named like \texttt{proc******\_Database} for each partition. These files will be needed for creating the distributed databases, and have to reside in the directory  \texttt{LOCAL\_PATH} specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}. Please see Chapter~\ref{cha:Creating-Distributed-Databases} for further details.
\end{itemize}
Note that all the files generated by the Python script  \texttt{cubit2specfem3d.py} must be placed in the \texttt{input\_directory} folder before running the program.

@@

After using \texttt{xmeshfem3D} or \texttt{xdecompose\_mesh}, the next step in the workflow is to compile
\texttt{xgenerate\_}~\\
\texttt{databases}. This program is going to create all the missing information needed by the SEM solver.

@@

the partitioner, i.e. from \texttt{xdecompose\_mesh} or \texttt{xmeshfem3D}.
